---
title: "Internal Engagement - Cyrus Herder"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: flatly
    source_code: embed
    social: ["linkedin","twitter","facebook"]
---

```{r setup, include=FALSE}
rm(list = ls())

path <- paste("/Volumes/DOCUMENTS/Personal/Advancement/")

knitr::opts_knit$set(root.dir = normalizePath(path))

pkgs <- c("dplyr","ggplot2","ggthemes","lubridate","kableExtra","flexdashboard","stringi","data.table",
          "tabulizer","readxl","xlsx","tidyr","treemap","forecast","viridisLite","AER","scales","highcharter","dygraphs")

miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] # vector of missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
invisible(lapply(pkgs,library,character.only=TRUE))

pal <- scale_fill_manual(values = c("#b2d234","#038ecd","#00377c","#a53b47","#4d4d4f",
                                    "#7798BF", "#AAEEEE"))


theme_hc <- function(){
  theme(
    text                = element_text(family="Source Sans Pro", size = 12),
    plot.title = element_text(family="Source Sans Pro Semibold", size = rel(1.4), 
                              hjust = 0.5),
    axis.line.x = element_line(colour = "black", size=0.5),
    axis.text.x = element_text(family = "Source Sans Pro", size = rel(1.0)),
    axis.text.y = element_text(family = "Source Sans Pro", size = rel(1.2)),
    axis.title = element_text(family = "Source Sans Pro", size = rel(1.0)),
    #title               = element_text(hjust=0), 
    axis.title.x        = element_text(hjust=.5),
    axis.title.y        = element_text(hjust=.5),
    panel.grid.major.y  = element_line(color='gray', size = .3),
    panel.grid.minor.y  = element_blank(),
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x  = element_blank(),
    panel.border        = element_blank(),
    panel.background    = element_blank(),
    legend.position     = "bottom",
    legend.title        = element_blank(),
    legend.direction = "horizontal"
  )
}

thm <- 
  hc_theme(
    colors = c("#b2d234","#038ecd","#00377c"),
    chart = list(
      backgroundColor = "transparent",
      style = list(fontFamily = "Source Sans Pro")
    ),
    xAxis = list(
      gridLineWidth = 1
    )
  )
```


Introduction {data-icon="fa-info-circle"}
===================================== 

How to Set up a knitted Rmarkdown in form of Flexdashboard

### Lung Deaths (All)

```{r}
dygraph(ldeaths)
```

### Lung Deaths (Male)

```{r}
dygraph(mdeaths)
```

### Lung Deaths (Female)

```{r}
dygraph(fdeaths)
```



Analysis {data-icon="fal fa-wifi"}
=====================================  

**Does the use of a contextually appropriate instrument affect that likelihood that Kibera residents rank as more likely a less probable but more 'representative' option in the 'Linda' problem?**



Column {data-width=600}
-----------------------------------------------------------------------

### Sales Forecast

```{r}
AirPassengers %>% 
  forecast(level = 90) %>% 
  hchart() %>% 
  hc_add_theme(thm)
```


### Sales by State

```{r}
data("USArrests", package = "datasets")
data("usgeojson")

USArrests <- USArrests %>%
  mutate(state = rownames(.))

n <- 4
colstops <- data.frame(
  q = 0:n/n,
  c = substring(viridis(n + 1), 0, 7)) %>%
  list_parse2()

highchart() %>%
  hc_add_series_map(usgeojson, USArrests, name = "Sales",
                    value = "Murder", joinBy = c("woename", "state"),
                    dataLabels = list(enabled = TRUE,
                                      format = '{point.properties.postalcode}')) %>%
  hc_colorAxis(stops = colstops) %>%
  hc_legend(valueDecimals = 0, valueSuffix = "%") %>%
  hc_mapNavigation(enabled = TRUE) %>%
  hc_add_theme(thm)
```

Column {.tabset data-width=400}
-----------------------------------------------------------------------

### Sales by Category

```{r, fig.keep='none'}
data("Groceries", package = "arules")
dfitems <- tbl_df(Groceries@itemInfo)

set.seed(10)

dfitemsg <- dfitems %>%
  mutate(category = gsub(" ", "-", level1),
         subcategory = gsub(" ", "-", level2)) %>%
  group_by(category, subcategory) %>% 
  summarise(sales = n() ^ 3 ) %>% 
  ungroup() %>% 
  sample_n(31)

tm <- treemap(dfitemsg, index = c("category", "subcategory"),
              vSize = "sales", vColor = "sales",
              type = "value", palette = rev(viridis(6)))

highchart() %>% 
  hc_add_series_treemap(tm, allowDrillToNode = TRUE,
                        layoutAlgorithm = "squarified") %>% 
  hc_add_theme(thm)
```

### Best Sellers

```{r}
set.seed(2)

nprods <- 10

dfitems %>% 
  sample_n(nprods) %>% 
  .$labels %>% 
  rep(times = sort(sample( 1e4:2e4, size = nprods), decreasing = TRUE)) %>% 
  factor(levels = unique(.)) %>% 
  hchart(showInLegend = FALSE, name = "Sales", pointWidth = 10) %>% 
  hc_add_theme(thm) %>% 
  hc_chart(type = "bar")
  
```


Overall Cases {data-navmenu="Special Cases" data-icon="fa-check"}
=====================================

**Distribution of Average Cases overall**

Row 
-------------------------------------
   
### Group Means - Overall
```{r}
RC<- USArrests%>%
  tidyr::gather(Types, Values,-state)%>%
  group_by(Types)%>%
  summarise(Mean_val = mean(Values))%>%
  ggplot(aes(x= reorder(Types,Mean_val), y= Mean_val))+
  geom_bar(stat = "identity", fill = "#038ecd")+
  geom_text(aes(label=Mean_val),vjust = -0.25, 
            size = 4, position = position_dodge(0.9),family="Source Sans Pro Semibold")+
  labs(y="Average",x="Type",title="Averages Cases Reported")+
  theme_hc()+pal
RC  
```


Cases by State {data-navmenu="Special Cases" data-icon="fa-check"}
=====================================

**Distribution of Average Cases overall**

Column {.tabset data-width=400}
-------------------------------------


### Murder >10

```{r}

ggplot(USArrests%>%filter(Murder>10),aes(x= reorder(state,Murder), y= Murder))+
  geom_bar(stat = "identity", fill = "#038ecd")+
  geom_text(aes(label=Murder),vjust = -0.25, 
            size = 4, position = position_dodge(0.9),family="Source Sans Pro Semibold")+
  labs(y="Average",x="",title="Averages Murder Cases Reported")+
  theme_hc()+ coord_flip()
```


Column
-------------------------------------

### Assault > 220

```{r}

ggplot(USArrests%>%filter(Assault>220),aes(x= reorder(state,Assault), y= Assault))+
  geom_bar(stat = "identity", fill = "#038ecd")+
  geom_text(aes(label=Assault),vjust = -0.25, 
            size = 3, position = position_dodge(0.9),family="Source Sans Pro Semibold")+
  labs(y="Average",x="",title="Averages Assault Cases Reported")+
  theme_hc()+ coord_flip()
```

### UrbanPop >75

```{r}

ggplot(USArrests%>%filter(UrbanPop>75),aes(x= reorder(state,UrbanPop), y= UrbanPop))+
  geom_bar(stat = "identity", fill = "#038ecd")+
  geom_text(aes(label=UrbanPop),vjust = -0.25, 
            size = 4, position = position_dodge(0.9),family="Source Sans Pro Semibold")+
  labs(y="Average",x="",title="Averages Urban Population Cases Reported")+
  theme_hc()+ coord_flip()
```

Values {data-icon="far fa-star-half-alt"}
===================================== 


Row
-------------------------------------


### Articles per Day

```{r}
computeArticles = function(...) return(45)
computeComments = function(...) return(126)
computeSpam = function(...) return(15)

articles <- computeArticles()
valueBox(articles, icon = "fa-pencil")
```

### Comments per Day

```{r}
comments <- computeComments()
valueBox(comments, icon = "fa-comments")
```



### Spam per Day

```{r}
spam <- computeSpam()
valueBox(spam, 
         icon = "fa-trash",
         color = ifelse(spam > 10, "warning", "primary"))
```

Row
-------------------------------------


### Linked Value Box

```{r}
valueBox(42, icon = "fa-pencil", href="#details")
```


Rating {data-icon="fas fa-door-open"}
===================================== 


Row {data-width=4}
-----------------------------------------------------------------------

### Contact Rate

```{r}
gauge(91, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)
))
```

### Average Rating 

```{r}
gauge(37.4, min = 0, max = 50, gaugeSectors(
  success = c(41, 50), warning = c(21, 40), danger = c(0, 20)
))
```

Row {data-width=4}
-----------------------------------------------------------------------

### Cancellations

```{r}
gauge(7, min = 0, max = 10, gaugeSectors(
  success = c(0, 2), warning = c(3, 6), danger = c(7, 10)
))
```


